Predicting Sequence Variability from Voronoi Tessellation of Proteins

### 2015-02-25 manuscript/voronoi/manuscript_voronoi.tex
MN update

    Motivated by the existing gaps in the current understanding of the role of flexibility and other structural properties on sequence-structure relations in proteins, here we propose and derive a new set of site-specific structural properties which unlike CN and WCN do not involve any adjustable free parameter in their definitions, while performing equally well or better than all previously-considered structural quantities in predicting protein sequence evolution. This is done by employing tessellation methods from the field of computational geometry to calculate several new site-specific quantities for proteins, which can serve as proxy measures of local packing density and site-specific flexibility. Contrary to what is currently perceived about the role of flexibility in sequence variability, we show that the newly calculated flexibility measures outperform many of previously studied structural properties, such as RSA and the traditional definitions of Contact Number and the Weighted Contact Number (WCN), in predicting sequence evolution at residue level. \\

    Furthermore, for structural properties that are calculated based on a set of representative site coordinates,  we show that the choice of the geometric average of the side chain atomic coordinates instead of the traditional choice of $C_\alpha$ atomic coordinates, always results in significantly better predictions of site-specific sequence evolution. We also show that the original kernel proposed for the definition of Weighted Contact Number by (xx) and supported further by (xx) and extensively used in other studies, has no significant advantage whatsoever in predicting B factors or the sequence variability, when compared to other possible types of kernels. A discussion of the methodology used in this work, the results and implications of our findings on the energy landscape of proteins and sequence-structure relations will be presented in the following sections.  \\

>>>>

    Motivated by the existing gaps in the current understanding of the role of flexibility and other structural properties on sequence-structure relations in proteins, here we propose and derive a new set of site-specific structural properties which, unlike CN and WCN, are not defined with any free parameters, while performing equally well or better than all previously-considered structural quantities in predicting protein sequence evolution. This is done by employing tessellation methods from the field of computational geometry to calculate several new site-specific quantities for proteins, which can serve as proxy measures of local packing density and site-specific flexibility. Contrary to what is currently perceived about the role of flexibility in sequence variability, we show that the newly calculated flexibility measures outperform many of previously studied structural properties, such as RSA and the traditional definitions of Contact Number and the Weighted Contact Number (WCN), in predicting sequence evolution at residue level. \\

    Furthermore, for structural properties that are calculated based on a set of representative site coordinates,  we show that the choice of the geometric average of the side chain atomic coordinates instead of the traditional choice of $C_\alpha$ atomic coordinates, always results in significantly better predictions of site-specific sequence evolution. Similar improvements in correlations with different ste-specific structural properties and sequence variability measures are also observed if the average of side chain B factors, instead of $C_\alpha$ atomic B factor, is used as a proxy measure of site flexibility.

    %We also show that the original kernel proposed for the definition of Weighted Contact Number by (xx) and supported further by (xx) and extensively used in other studies, has no significant advantage whatsoever in predicting B factors or the sequence variability, when compared to other possible types of kernels. A discussion of the methodology used in this work, the results and implications of our findings on the energy landscape of proteins and sequence-structure relations will be presented in the following sections.  \\
